Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex
spelling

The studies mentioned before all depend on the question of whether fMRI data carry class
information (pattern discrimination) and exploiting this information to perform classification. However, researchers have indicated at least two other basic questions to which classifiers can be applied, being ``where or when
is class information represented?'' (pattern localization) and ``how is
class information encoded'' or ``how does its encoding relate to known
relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}. While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.\\

@@

\section{Methods}
\label{sec:methods}
We now continue by discussing the methods applied in our research. The general approach taken for each of the algorithms discussed consists of the following steps:
\begin{enumerate}
  \item Preprocessing
  \item Dimensionality reduction
  \item Training
  \item Testing
\end{enumerate}
Preprocessing was done to transform the dataset into a format suitable for input into the classification algorithm. Dimensionality reduction was a direct result of the large dimensionality of the dataset and may consist of both feature extraction as well as feature selection. Finally, the approach taken to training and testing the classifier deals with preventing overfitting the classifier to training data and verifying accuracy for cross-subject classification. The choices made in each of these steps may affect classifier accuracy, so to obain a complete understanding of classification accuracies we have tried several methods at each step. In the following sections, these methods will be described in more detail.\\
The implementation of the algorithms was done using Python combined with modules tailored to mathematics (NumPy, SciPy) and machine learning (ScikitLearn). The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{preprocessing}). The source code for the algorithms is available from the authors on request.

\subsection{Data Acquisition and Dimensions}
The dataset used for our experiments was originally collected by Marcel Just et al. from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset. This set consists of fMRI data for 6 subjects that was obtained while having each subject participate in a number of trials.

>>>>

The studies mentioned before all depend on the question of whether fMRI data carry class
information (pattern discrimination) and exploiting this information to perform classification. However, researchers have indicated at least two other basic questions to which classifiers can be applied, being ``where or when
is class information represented?'' (pattern localization) and ``how is
class information encoded'' or ``how does its encoding relate to known
relationships between stimuli'' (pattern characterization)~\cite{pereira2009machine}. While the main focus of this paper is on the application of machine learning techniques in pattern discrimination, we will briefly discuss these two other applications.

@@

\section{Methods}
\label{sec:methods}
We now continue by discussing the methods applied in our research. The general approach taken for each of the algorithms discussed consists of the following steps:
\begin{enumerate}
  \item Preprocessing
  \item Dimensionality reduction
  \item Training
  \item Testing
\end{enumerate}
Preprocessing was done to transform the dataset into a format suitable for input into the classification algorithm. Dimensionality reduction was required to cope with the high dimensionality of the dataset and may consist of both feature extraction as well as feature selection. Finally, the approach taken to training and testing the classifier deals with preventing overfitting the classifier to training data and verifying accuracy for cross-subject classification. The choices made in each of these steps may affect classifier accuracy, so to obain a complete understanding of classification accuracies we have tried several methods at each step. In the following sections, these methods will be described in more detail.\\
The implementation of the algorithms was done using Python combined with modules tailored to mathematics (NumPy, SciPy) and machine learning (ScikitLearn). The original dataset was made available as a MatLab file and we used preprocessing in MatLab to create a number of files containing modified versions of this data, such as normalization and ROI grouping (see section~\ref{preprocessing}). The source code for the algorithms is available from the authors on request.

\subsection{Data Acquisition and Dimensions}
The dataset used for our experiments was originally collected by Marcel Just and colleagues from Carnegie Mellon University's Center for Cognitive Brain Imaging and published as the StarPlus fMRI dataset. This set consists of fMRI data for 6 subjects that was obtained while having each subject participate in a number of trials.
